\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsthm} %numéroter les questions
\usepackage[english]{babel}
\usepackage{datetime}
\usepackage{xspace} % typographie IN
\usepackage{hyperref}% hyperliens
\usepackage[all]{hypcap} %lien pointe en haut des figures
%\usepackage[french]{varioref} %voir x p y

\usepackage{fancyhdr}% en têtes
\usepackage{graphicx} %include pictures
\usepackage{pgfplots}

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{babel}
\usepackage{circuitikz}
% \usepackage{gnuplottex}
\usepackage{float}
\usepackage{ifthen}

\usepackage[top=1.3 in, bottom=1.3 in, left=1.3 in, right=1.3 in]{geometry}
\usepackage[]{pdfpages}
\usepackage[]{attachfile}

\usepackage{amsmath}
\usepackage{enumitem}
\setlist[enumerate]{label=\alph*)}% If you want only the x-th level to use this format, use '[enumerate,x]'
\usepackage{multirow}

\usepackage{aeguill} %guillemets


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% READ THIS BEFORE CHANGING ANYTHING
%
%
% - TP number: can be changed using the command
%		\def \tpnumber {TP 1 }
%
% - Version: controlled by a new command:
%		\newcommand{\version}{v1.0.0}
%
% - Booleans: there are three booleans used in this document:
%	- 'corrige', controlled by defining the variable 'corrige'
% You can define those variables in a makefile using such a command:
% pdflatex -shell-escape -jobname="infoh410_tp1" "\def\corrige{} \input{infoh410_tp1.tex}"
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Numero du TP :
\def \tpnumber {TP 4 }

\newcommand{\version}{v1.0.0}


% ########     #####      #####    ##         #########     ###     ##      ##  
% ##     ##  ##     ##  ##     ##  ##         ##           ## ##    ###     ##  
% ##     ##  ##     ##  ##     ##  ##         ##          ##   ##   ## ##   ##  
% ########   ##     ##  ##     ##  ##         ######     ##     ##  ##  ##  ##  
% ##     ##  ##     ##  ##     ##  ##         ##         #########  ##   ## ##  
% ##     ##  ##     ##  ##     ##  ##         ##         ##     ##  ##     ###  
% ########     #####      #####    #########  #########  ##     ##  ##      ##  
\newboolean{corrige}
\ifx\correction\undefined
\setboolean{corrige}{false}% pas de corrigé
\else
\setboolean{corrige}{true}%corrigé
\fi
\setboolean{corrige}{false}% pas de corrigé


\definecolor{darkblue}{rgb}{0,0,0.5}

\pdfinfo{
/Author (ULB -- CoDe/IRIDIA)
/Title (\tpnumber, INFO-H-410)
/ModDate (D:\pdfdate)
}

\hypersetup{
pdftitle={\tpnumber [INFO-H-410] Techniques of AI},
pdfauthor={ULB -- CoDe/IRIDIA},
pdfsubject={}
}

\theoremstyle{definition}% questions pas en italique
\newtheorem{Q}{Question}[] % numéroter les questions [section] ou non []


%  ######     #####    ##       ##  ##       ##     ###     ##      ##  ######      #######   
% ##     ##  ##     ##  ###     ###  ###     ###    ## ##    ###     ##  ##    ##   ##     ##  
% ##         ##     ##  ## ## ## ##  ## ## ## ##   ##   ##   ## ##   ##  ##     ##  ##         
% ##         ##     ##  ##  ###  ##  ##  ###  ##  ##     ##  ##  ##  ##  ##     ##   #######   
% ##         ##     ##  ##       ##  ##       ##  #########  ##   ## ##  ##     ##         ##  
% ##     ##  ##     ##  ##       ##  ##       ##  ##     ##  ##     ###  ##    ##   ##     ##  
%  #######     #####    ##       ##  ##       ##  ##     ##  ##      ##  ######      #######   
\newcommand{\reponse}[1]{% pour intégrer une réponse : \reponse{texte} : sera inclus si \boolean{corrige}
\ifthenelse {\boolean{corrige}} {\fr{\paragraph{Réponse :}}\en{\paragraph{Answer:}} \color{darkblue}   #1\color{black}} {}
}

\newcommand{\addcontentslinenono}[4]{\addtocontents{#1}{\protect\contentsline{#2}{#3}{#4}{}}}


%  #######   ##########  ##    ##   ##         #########  
% ##     ##      ##       ##  ##    ##         ##         
% ##             ##        ####     ##         ##         
%  #######       ##         ##      ##         ######     
%        ##      ##         ##      ##         ##         
% ##     ##      ##         ##      ##         ##         
%  #######       ##         ##      #########  #########  
%% fancy header & foot
\pagestyle{fancy}
\lhead{[INFO-H-410] Techniques of AI\\ \tpnumber}
\rhead{\version\\ page \thepage}
\chead{\ifthenelse{\boolean{corrige}}{Correction}{}}
\cfoot{}
%%

\setlength{\parskip}{0.2cm plus2mm minus1mm} %espacement entre §
\setlength{\parindent}{0pt}

% ##########  ########   ##########  ##         #########  
%     ##         ##          ##      ##         ##         
%     ##         ##          ##      ##         ##         
%     ##         ##          ##      ##         ######     
%     ##         ##          ##      ##         ##         
%     ##         ##          ##      ##         ##         
%     ##      ########       ##      #########  ######### 
\date{\vspace{-1.7cm}\version}
\title{\vspace{-2cm} \tpnumber \\ Techniques of AI [INFO-H-410] \ifthenelse{\boolean{corrige}}{Correction}{}}


\begin{document}
\selectlanguage{english}

\maketitle

\paragraph{Representation and Interpretation of Boolean Functions}
\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Symbol}            & \textbf{Name} \\ \hline
0                          & FALSE         \\ \hline
1                          & TRUE          \\ \hline
$!A$ / $\neg A$ & NOT A         \\ \hline
$A \wedge B$               & A AND B \\ \hline
$A \vee B$                 & A OR B \\ \hline
$A \oplus B$                 & A XOR B \\ \hline
\end{tabular}
\end{center}

\paragraph{Decision Trees}
\begin{Q}
Give decision trees to represent the following Boolean functions:
\begin{enumerate}
    \item $A \wedge \neg B$
    \item $A \vee (B \wedge C)$
    \item $A \oplus B$
    \item $(A \wedge B) \vee (C \wedge D) $
\end{enumerate}
\reponse{
}
\end{Q}

\begin{Q}
True or false: any Boolean function can be expressed as a decision tree. Give a proof.
\reponse{
}
\end{Q}


\begin{Q}
True or False: If a decision tree D2 is an elaboration of tree D1, then D1 is more general than
D2. Assume D1 and D2 are decision trees representing arbitrary Boolean functions, and that D2
is an elaboration of D1 if ID3 could extend D1 into D2. If true, give a proof; if false, give a
counter example.

More-general-than was defined in chapter two: Let $h_j$ and $h_k$ be Boolean-valued functions defined
over $X$. Then $h_j$ is more general than or equal to $h_k$ if and only if 
$(\forall x \in X)$ $h_k(x) = 1 \rightarrow h_j(x) = 1 $

\reponse{
}
\end{Q}

\paragraph{ID3}
\begin{Q}
In order to evaluate the quality of the tree which is grown by ID3, one could compare its performance
to a baseline performance. The baseline performance is often the performance of a very simple
machine learning algorithm. Consider the following approach:


You have a dataset consisting of 25 examples of two classes. You plan to use leave-one-out cross
validation. As a baseline, you use a simple majority classifier (a majority classifier is given a set of
training data and then always outputs the class that is in the majority in the training set, regardless
the input). Such a majority classifier is expected to score about 50\%, but with this example of
leave-one-out cross-validation, it does not. What will be its performance and why?
\end{Q}


\begin{Q}
Consider the following data on the hair color, body weight, body height and the usage of lotion of
eight different people. The table shows whether the people got sunburned after an afternoon in the
sun.
% TODO : add table
\begin{enumerate}
    \item Perform average entropy calculations on the following complete dataset for each of the four
    attributes. Select the attribute which minimizes the entropy; draw the first level of the decision
    tree.
    \item Grow the tree until you reach the proper identification of all the samples.
    \item Establish the rules from the tree found in the previous question.
    \item The factual value of the training instance, in the dataset is: \textbf{Sunburned? – Yes}. 
    \end{enumerate}
% TODO : add table
        Consider the following additional instance in your dataset:
Will the inclusion of this new instance in your training dataset impact the structure of the
decision tree? What are such instances considered to be? Does their presence or absence in
the training dataset impact the outcome of the built model?
\end{Q}

\begin{Q}
    Consider the herewith provided alternative dataset:
    % TODO : add table
    Is this dataset suitable for training your model and will you be able to build an optimal decision
    tree? Explain briefly how you can deal with this situation.
\end{Q}

\begin{Q}
    Consider the data set given below. Explain how one can handle a continuous-valued attribute when
building a decision tree and demonstrate the approach on the given data.
    % TODO : add table
\end{Q}




\end{document} 
